---
title: "Sampling error and bootstrapping"
output: html_notebook
---

```{r}
source("constants_and_functions.R")
```

How close are estimations of sampling error through bootstrapping?

```{r}
general_sample_size = 10**4
p_a = 0.4
p_b = 0.4


layout(matrix(c(1,2), ncol = 2, byrow = TRUE))
general_vector_of_haplotypes <- get_vector_of_haplotypes(number_of_individuals = general_sample_size, 
                                                         p_a = p_a, p_b = p_b, r = 0)
r_values <- replicate(10000,  get_r_from_haplotypes(sample(general_vector_of_haplotypes, 200)))
hist(r_values, xlab = 'r', main = '',breaks = 20)
abline(v=get_r_from_haplotypes(general_vector_of_haplotypes), col='blue')

general_vector_of_haplotypes <- sample(haplotype_constants, size=general_sample_size, replace=TRUE, prob=get_haplotypes_freq(p_a = p_a, p_b = p_b, r = 0.9))
r_values <- replicate(10000, get_r_from_haplotypes(sample(general_vector_of_haplotypes, 200)))
hist(r_values, xlab = 'r', main='', breaks = 20)
abline(v=get_r_from_haplotypes(general_vector_of_haplotypes), col='blue')
print(mean(r_values))

```

```{r}
#bootstrapping

small_sample_size = 200

layout(matrix(c(1,2), ncol = 2, byrow = TRUE))
small_sample = sample(haplotype_constants, size=small_sample_size, replace=TRUE, prob=get_haplotypes_freq(p_a = p_a, p_b = p_b, r = 0))
results <- boot(data=small_sample, statistic=LD, R=1000)
hist(results$t, breaks = 20, xlab = 'r', main='r_real=0.0')
abline(v=results$t0, col='blue')
small_sample = sample(haplotype_constants, size=small_sample_size, replace=TRUE, prob=get_haplotypes_freq(p_a=p_a, p_b=p_b, r=0.9))
results <- boot(data=small_sample, statistic=LD, R=1000)
hist(results$t, breaks = 20, xlab = 'r', main='r_real=0.9')
abline(v=results$t0, col='blue')
```

How LD calculated from genotypes and haplotypes match?

```{r}
get_r_genotype_haplotype <- function(sample_size, p_a, p_b, r){
  haplotypes <- get_vector_of_haplotypes(number_of_individuals=sample_size, p_a=p_a, p_b=p_b, r=r)
  c(get_r_from_genotypes(haplotypes), get_r_from_haplotypes(haplotypes))
}

rs <- as.matrix(replicate(1000, get_r_genotype_haplotype(sample_size = 100, 0.5,0.5, r=0.8)), nrow =2)
plot(x = rs[1,], y = rs[2,], xlab = "r from genotypes", ylab = "r from haplotypes")
```
We can see that estimation of corrlation via genotypes causes larger sampling error.

```{r}
get_sampling_error_for_specific_parameters <- function(sample_size, p_a, p_b, r, number_of_simulations){
  rs <- as.matrix(replicate(number_of_simulations, get_r_genotype_haplotype(sample_size = sample_size, p_a=p_a,p_b=p_b, r=r)), nrow =2)
  c(sd(rs[1,]), sd(rs[2,]))
}

standard_deviation <- as.matrix(replicate(200, get_sampling_error_for_specific_parameters(sample_size = 100, 0.5,0.5, r=runif(1), number_of_simulations=200)), nrow =2)
plot(x = standard_deviation[1,], y = standard_deviation[2,], xlab = "sd from genotypes", ylab = "sd from haplotypes")
abline(b=1, a =0)


```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).
